\input{chaselev/macros}

\section{Chase-Lev work-stealing deque}
\label{sec:chaselev}

\input{chaselev/figures/work_stealing}

\paragraph{Work-stealing.}

Randomized \emph{work stealing}~\citep{DBLP:journals/jacm/BlumofeL99} is the standard strategy for parallel task scheduling.
It has been implemented in many libraries, including \Cilk~\citep{DBLP:journals/jpdc/BlumofeJKLRZ96,DBLP:conf/pldi/FrigoLR98}, \TBB, \OpenMP, \Taskflow~\citep{DBLP:journals/tpds/HuangLLL22}, \Tokio and \Domainslib~\citep{domainslib}.

\paragraph{Work-stealing deque.}

The idea of work-stealing, illustrated in \cref{fig:chaselev_work_stealing}, is the following.
Each domain owns a deque-like data structure, called \emph{work-stealing deque}, to store its tasks.
Locally, each domain treats its deque as a stack, operating at the back end.
When a domain runs out of tasks, it becomes a thief: it tries to steal a task from the deque of another randomly selected ``victim'' domain, operating at the front end.
Multiple thieves may concurrently attempt to steal tasks from a single deque.

The fact that \ocamlinline{steal} works on the opposite end from \ocamlinline{push/pop} has two purposes. First, the owner domain and a thief more rarely race on the same task. Second, the owner domain pops tasks and pushes smaller sub-tasks, so tasks are typically smaller at the back: the owner pops small tasks that it pushed recently, good for locality, while thieves steal larger tasks, making steals rarer.

\paragraph{Deque implementations.}

The most popular work-stealing deque algorithm is the Chase-Lev deque~\citep{DBLP:conf/spaa/ChaseL05,DBLP:conf/ppopp/LePCN13}; it is lock-free and unbounded.
We verified the implementation from the \Saturn library~\citep{saturn}~\refLibTheoriesFile{zoo_saturn/ws_deque_1} along with two other variants: a bounded variant~\refLibTheoriesFile{zoo_saturn/ws_bdeque_1}, used in the \Moonpool~\citep{moonpool} and \Taskflow~\citep{DBLP:journals/tpds/HuangLLL22} libraries, and an idealized infinite-array-based variant~\refLibTheoriesFile{zoo_saturn/inf_ws_deque_1}.

Remarkably, the three variants essentially share the same logical states.
In particular, although they do not behave exactly the same way, the original and the idealized versions follow a similar concurrent protocol, involving external and future-dependent linearization.

\subsection{Interface}

The three \ocamlinline{deque} implementations we describe below share the same programming interface \refLib{zoo_saturn/ws_deque_2.mli}.

\ocamlinline{create} returns a new queue:
\begin{description}
\item[\ocamlinline{val create : unit -> 'a t}]
\end{description}

``Private'' operations may only be called by the domain owning the deque:
\begin{description}
\item[\ocamlinline{val size : 'a t -> int}] returns the number of elements on the deques.

\item[\ocamlinline{val is_empty : 'a t -> bool}] checks if the queue is empty.

\item[\ocamlinline{val push : 'a t -> 'a -> unit}] pushes an element to the ``back'' the queue.

\item[\ocamlinline{val pop : 'a t -> 'a option}] pops an element from the the ``back'' of the queue.
\end{description}

Finally \ocamlinline{steal} is ``public'', it may be called by any domain:
\begin{description}
\item[\ocamlinline{val steal : 'a t -> 'a option}] attempts to steal
  an element from the ``front'' of the queue.
\end{description}

\input{chaselev/inf_ws_deque}

\input{chaselev/ws_bdeque}

\input{chaselev/ws_deque}
