\section{Futures}
\label{sec:parabs_future}

At the fourth level, \ocamlinline{Future}~\refLibTheoriesFile{zoo_parabs/future} implements futures%
\footnote{
Futures are called \emph{promises} in \Domainslib.
In fact, the two notions are often used in conjunction to represent the two sides of the same object.
}%
, a standard abstraction for representing the future result of an asynchronous task.

\subsection{Interface}
\label{sec:parabs_future_intf}

The OCaml interface of the \ocamlinline{Future} module is as follows~\refLib{zoo_parabs/future.mli}:

\begin{description}
\item[\ocamlinline{type 'a t}] represents the future result of an asynchronous task,
  of type \ocamlinline{'a}.

\item[\ocamlinline{val return : 'a -> 'a t}] returns the result of
  trivial task that is already finished.

\item[\ocamlinline{val async : Pool.context -> 'a Pool.task -> 'a t}] schedules the given
  task to be executed asynchronously, and returns a future for its result.

\item[\ocamlinline{val wait : Pool.context -> 'a t -> 'a}] blocks
  until the result of the given future is available and returns it;
  the caller domain helps running scheduler tasks in the meantime.

\item[\ocamlinline{val iter : Pool.context -> 'a t -> ('a -> unit)
    Pool.task -> unit}] schedules a unit-returning task to run after
  the result of a future is available, and returns immediately.

\item[\ocamlinline{val map : Pool.context -> 'a t -> ('a -> 'b)
    Pool.task -> 'b t}] schedules a task to run after the result of
  a future is available, and returns a future for its result.
\end{description}

\paragraph{Example}
We can now write a higher-level version of the
\ocamlinline{fibo} function from \cref{sec:parabs_pool_intf}:
\begin{ocamlcode}
let rec fibo ctx n =
  if n <= sequential_cutoff then fibo_seq n
  else
    let a = Future.async ctx (fun ctx -> fibo ctx (n - 1)) in
    let b = fibo ctx (n - 2) in
    Future.wait ctx a + b
\end{ocamlcode}

\subsection{Specification}
\label{sec:parabs_future_spec}

\begin{tirPrefix}{future-}

\input{parabs/figures/future_spec}

The specification is given in \cref{fig:parabs_future_spec}.
It features four predicates: \inv, \result, \consumer and \obligation.

\c{async} allows submitting a task asynchronously while executing under a context (\refTirName{sync-spec}), returning a \emph{future} representing the result of the task.
To actually get the result, one must call \c{wait} (\refTirName{wait-spec}).
$\c{iter}\ \v{ctx}\ \v{fut}\ \v{task}$ attaches callback \v{task} to \v{fut} (\refTirName{iter-spec}) and $\c{map}\ \v{ctx}\ \v{fut}_1\ \v{task}$ creates a new future to be resolved after $\v{fut}_1$ (\refTirName{map-spec}).

The persistent assertion $\inv\ \v{pool}\ t\ \iPredTwo\ \iPredFour$ represents the knowledge that $t$ is a valid future attached to pool \v{pool} such that:
(1) \iPredTwo is the \emph{non-persistent output predicate} satisfied by the produced value;
(2) \iPredFour is the \emph{persistent output predicate} satisfied by the produced value.

The persistent assertion $\result\ t\ \zooVal$ represents the knowledge that future $t$ has been resolved to value \zooVal.
Using \refTirName{inv-result}, it can also be combined with \inv to obtain the persistent output predicate.
After the pool has finished, it is guaranteed that all futures have been resolved (\refTirName{inv-finished}).

The assertion $\consumer\ t\ \iPredThree$ represents the right to consume \iPredThree once future $t$ has been resolved.
Indeed, using \refTirName{inv-result-consumer}, it can be combined with \inv and \result to obtain \iPredThree.
When $t$ is created, this assertion is produced with the full non-persistent predicate (\refTirName{async-spec}, \refTirName{map-spec}); then, it can be divided into several parts (\refTirName{consumer-divide}).

The persistent assertion $\obligation\ \v{pool}\ \iProp$ represents a proof obligation emitted by \c{iter} (\refTirName{iter-spec}).
It allows retrieving \iProp once \v{pool} has finished (\refTirName{obligation-finished}).

One notable aspect of this specification is that resolution of the future --- as indicated by \result\xspace --- is separated from the division of the output predicates --- as achieved by \consumer.

\end{tirPrefix}

\subsection{Implementation}
\label{sec:parabs_future_impl}

Futures are implemented using \emph{ivars} (concurrent write-once variables), as implemented and verified in the \Zoo standard library.
\c{async} creates an ivar and calls \ocamlinline{Pool.async} to resolve it asynchronously.
\c{wait} calls \ocamlinline{Pool.wait_until} to wait \emph{actively} until the ivar is resolved and returns the resulting value.
