\section{Pool}
\label{sec:parabs_pool}

At the third level, \ocamlinline{Pool}~\refLibTheoriesFile{zoo_parabs/pool} implements a task scheduler on top of a given realization of \ocamlinline{Ws_hub}.
It offers essentially the same functionalities as \Domainslib with a few notable differences.
(1) Exceptions raised by tasks are not caught and therefore not re-raised properly by the scheduler since \ZooLang does not currently support them.
(2) Since \ZooLang does not support algebraic effects~\citep{DBLP:conf/pldi/Sivaramakrishnan21} either, the interface is slightly more involved (see \emph{execution contexts} in \cref{sec:parabs_pool_spec}).

Moreover, this limitation imposes a \emph{child-stealing} strategy, as opposed to a \emph{continuation-stealing} strategy that would require capturing the continuation of a computation.

Also, this makes it difficult to implement a \c{yield} operation%
\footnote{
\Domainslib does not currently provide a \c{yield} operation but it can be easily implemented.
}%
, \ie an operation that yields control to the scheduler, letting it reschedule the current task later.

\subsection{Interface}
\label{sec:parabs_pool_interface}

The \OCaml interface of the \ocamlinline{Pool} module is as follows~\refLib{zoo_parabs/pool.mli}:

\begin{description}
\item[\ocamlinline{type t}] represents a scheduler that owns a pool of
  $n$ worker domains and a set of tasks to compute. One domain owns
  the scheduler and can send toplevel tasks, so in total $n + 1$ domains
  participate to the computation.

\item[\ocamlinline{type context}] represents the runtime execution
  state of a running scheduler. All asynchronous functions take
  a \ocamlinline{context} parameter, which is passed to asynchronous
  operations.

\item[\ocamlinline{type 'a task = context -> 'a}]
\end{description}

\noindent
The following functions are the high-level part of the module,
needed to create a scheduler and run a toplevel task.
\begin{description}
\item[\ocamlinline{val create : int -> t}] creates a scheduler with
  the given number of worker domains.

\item[\ocamlinline{val run : t -> 'a task -> 'a}] runs a task
  on a scheduler. The call returns when the task is finished.

\item[\ocamlinline{val kill : t -> unit}] stops a scheduler,
  terminating workers when they become idle. This should be called
  once you know that the toplevel task is finished.
\end{description}

\noindent
The following low-level functions are typically not used directly, as
users would prefer the higher-level interface of the
\ocamlinline{Future} (\cref{sec:parabs_future}) or \ocamlinline{Algo}
(\cref{sec:parabs_algorithms}) modules to define asynchronous tasks.
\begin{description}
\item[\ocamlinline{val size : context -> int}] returns the number of
  worker domains, which can help when deciding how to split work in an
  appropriate number of asynchronous tasks.

\item[\ocamlinline{val async : context -> unit task -> unit}] schedules
  a given task to be executed asynchronously by a running scheduler.

\item[\ocamlinline{val wait_until : context -> (unit -> bool) -> unit}]
  waits until the provided predicate returns \ocamlinline{true},
  executing other tasks in the meantime.
\end{description}

\subsection{Example}
\label{sec:parabs_pool_example}

Suppose we have a
%
\ocamlinline{fibo : Pool.context -> int -> int}
%
function that computes fibonacci numbers naively but in parallel. Then
the following main expression will compute and print the $40$-th
fibonacci number.
\begin{ocamlcode}
let () =
  let pool = Pool.create (Domain.recommended_domain_count () - 1) in
  let fib40 = Pool.run pool (fun ctx -> fibo ctx 40) in
  Pool.kill pool;
  Printf.printf "fib 40 = %d\n%!" fib40
\end{ocamlcode}

The \ocamlinline{fibo} itself could be defined from the low-level
interface directly as follows (assume \ocamlinline{fibo_seq} is
a sequential definition of the fibonacci function):
\begin{ocamlcode}
let rec fibo ctx n =
  if n <= sequential_cutoff then fibo_seq n
  else
    let a = Atomic.make None in
    Pool.async ctx (fun ctx -> Atomic.set a (Some (fibo ctx (n - 1))));
    let b = fibo ctx (n - 2) in
    Pool.wait_until ctx (fun () -> Option.is_some (Atomic.get a));
    Option.get (Atomic.get a) + b
\end{ocamlcode}
Because \ocamlinline{Pool.async} does not wait on a result, we use an
atomic reference to store the result and wait on it. This pattern is
wrapped in an \ocamlinline{Ivar} module we contributed to the Zoo
standard library, but it is even better to use \ocamlinline{Future}
directly.

Note that the \ocamlinline{fibo} function takes a parameter \ocamlinline{ctx : Pool.context}. This is idiomatic, all user functions using \ocamlinline{Parabs} for asynchronous computations should take a context parameter. (The higher-level API \ocamlinline{Future} and \ocamlinline{Algo} also require contexts.) This is common practice of library-based schedulers: \Domainslib requires a \ocamlinline{pool} parameter and \Taskflow requires a \texttt{tf::Runtime\&} or \texttt{tf::Executor\&}.

\subsection{Specification}
\label{sec:parabs_pool_spec}

\begin{tirPrefix}{pool-}

\input{parabs/figures/pool_spec}

The specification is given in \cref{fig:parabs_pool_spec}.
It features five predicates: \inv, \model, \context, \finished and \obligation.

The persistent assertion $\inv\ t\ \v{sz}$ represents the knowledge that $t$ is a valid scheduler; \v{sz} is the number of worker domains.
It is returned by \c{create} (\refTirName{create-spec}) and required only by \c{size} (\refTirName{size-spec}).
Its only purpose is to record the immutable characteristics of the scheduler.

The assertion $\model\ t$ represents the ownership of scheduler $t$.
It is returned by \c{create} (\refTirName{create-spec}) and required by external operations (\refTirName{run-spec}, \refTirName{kill-spec}).
For example, $\c{run}\ t\ \v{task}$ submits \v{task} to scheduler $t$; it returns both \model and the output predicate of \v{task}.

The assertion $\context\ t\ \v{ctx}\ \v{scope}$ represents the ownership of \emph{execution context} \v{ctx} attached to scheduler $t$; \v{scope} is a purely logical parameter connecting input and output \context, which is necessary in the proof.
Any task execution happens under such a context (\refTirName{run-spec}, \refTirName{async-spec}, \refTirName{wait-until-spec}).
In particular, all internal operations require and return \context.
For example, $\c{async}\ \v{ctx}\ \v{task}$ submits \v{task} asynchronously while executing under context \v{ctx}; \v{task} must be shown to execute safely under any context attached to the same scheduler (\refTirName{async-spec}).

The persistent assertion $\finished\ t$ represents the knowledge that scheduler $t$ has finished, meaning all submitted tasks were executed.
It can be obtained by calling \c{kill} (\refTirName{kill-spec}).

The persistent assertion $\obligation\ t\ \iProp$ represents a proof obligation attached to scheduler $t$.
It allows retrieving \iProp once $t$ has finished executing (\refTirName{obligation-finished}).
Obligations are obtained by submitting tasks through \c{async} (\refTirName{async-spec}).

The assertion $\consumer\ t\ \iProp$ represents the right to consume \iProp once pool $t$ has finished executing (\refTirName{consumer-finished}).
Similarly to \obligation, it can be obtained through \c{async} (\refTirName{async-spec}).
Contrary to \obligation, it is non-persistent and can only be used once through \refTirName{consumer-finished}, which is sometimes less convenient.
A simple example of use is the parallel ``quicksort'' algorithm~\refLibTheoriesFile{examples/pool_quicksort}, where multiple nested \consumer assertions are generated.

\end{tirPrefix}

\subsection{Implementation}
\label{sec:parabs_pool_impl}

\paragraph{Worker domains.}

The implementation relies on a pool of worker domains and a work-stealing hub.
Each worker runs the following loop:
(1) get a task using \ocamlinline{Ws_hub.pop_steal};
(2) if it fails, the scheduler has been killed and so the worker stops, otherwise execute the task in the context of the current worker;
(3) start over.

\paragraph{Blocking.}

Care must be taken to block and unblock work-stealing deques properly.
When the scheduler is killed, it is crucial that workers block their deque before stopping; otherwise, the scheduler may never terminate because of a running worker waiting forever for a response from a stopped but unblocked worker.
Also, the main domain, from which tasks can be submitted externally through \c{run}, must unblock when it is executing tasks and block when it is not.

\paragraph{Awaiting.}

When \c{wait\_until} is called and the predicate we are waiting on does not hold, it executes a task from the queue before retrying recursively. In other words, for a given worker the tasks that are currently waiting are stored on the call stack, and they will be re-checked in LIFO order after the current task terminates.

Note: It would be nicer to store waiting tasks in a scheduler data structure; instead of a 'pull' design where waiting tasks check the future result regularly, one could 'push' tasks which become ready when the future they depend on is resolved. \Domainslib leverages algebraic effects for this: awaiting a future captures the continuation and stores it into the future; when the future is resolved, it resubmits all the waiting tasks.

\paragraph{Shutdown.}

In \Domainslib, scheduler shutdown consists in submitting special tasks through the main domain; when a worker finds such a task, it quickly stops.
However, this simple mechanism has at least two drawbacks:
(1) it introduces an indirection for every regular task, which may be expensive;
(2) it works well under standard work-stealing but is more difficult to implement under other scheduling strategies, especially work-stealing with private deques (see \cref{sec:parabs_ws_deques_private}).
Consequently, we use an alternative mechanism implemented at the level of \ocamlinline{Ws_hub}: a shared flag, regularly checked in \ocamlinline{Ws_hub.steal} and \ocamlinline{Ws_hub.pop_steal}, is set when the scheduler is killed.
